Multi-task learning in Computer Vision

### 2020-05-31 __latexindent_temp.tex
Final version. Hopefully it is not terrible :^).

The ibean dataset \citep{beansdata} contains 1296 images classified into three classes; Healthy, Angular Leaf Spot, and Bean Rust.
The data set provides a split for the images that will be the split that is used for the experiments.
The citrus leaves dataset \citep{citrusdata}, on the other hand, contains 759 examples of citrus fruits and leaves with six different classes; Black Spot, Canker, Greening, Scab, Melanose, and Healthy.
This set is quite tricky as it contains 609 images of leaves and only 150 images of actual citrus fruits, and some of the classes have very few images.
Some examples from the data sets are in Figure 7.1.
For the training split, the data will be split in 65/10/25 ratio for the training, validation, and test sets, respectively.
The goal of this kind of split is to have some relatively reliable test results by having proportionally larger than regular test set since there is so little data.
Both of these datasets have a relatively small amount of images available for training, so it should serve as a good basis for trying to train multi-task classifiers and give a glimpse into the adjustments that need to be made when training multi-task classifiers.
The goal for this experiment is to get familiar with the basics of how to create the multi-task models, and how to do the actual changes to the training loop in the Pytorch code to accommodate for the multi-task training.

When the two closely related datasets are used to train the model, we could expect the model to generalize better on both of them, as hopefully similar features would benefit both tasks.
But as the tasks are quite simple, the benefits can be relatively small as the likely accuracy for the single-task models should be quite high.
In the case of the citrus data set, learning to classify the citrus fruits could prove quite difficult, and for that, the bean leaves likely won't be of much help.
This model is depicted in figure 7.2.

@@

Training both tasks in a multi-task setting, we found that the accuracy on the citrus data set was increased, and ibean accuracy stayed the same.
Specifically, the accuracy for the EfficientNet classifier went up to 88\% and for ResNet up to 89\%.
The final model evaluated on the test set was the model with the highest average validation accuracy over the two tasks when trained for 25 epochs.
All the results are displayed in table 7.1.
The results show that even though these tasks were relatively simple and the accuracies for the single-task models were high, there was a benefit to be gained from multi-task training.
Here we mainly focused on getting good results on a fully shared backbone, and more experiments could be done by having fewer parameters shared between the tasks.
As an introduction to multi-task learning, the experiments show that picking the correct sampling ratio is extremely important for achieving desired results.
In that way, the sampling ratio seems quite unlike learning rate in that picking too small learning rate may lead to slow convergence but will most likely produce results in the end.
With the sampling ratio, it seems that the correct value needs to be found by experimentation, as a too low ratio for the citrus data gives accuracies that can be as low as 60\%.

@@

The second way of approaching this problem is to first train a model on the COCO classes and to create a detector for the cat and dog classes.
Using the high-level model, we can automatically create acceptable object detection labels for the 37 classes in the pets dataset.
Then we can initialize a 37 class object detector using the weights of the two-class model.
By having such a well-initialized model for the detection, the second part of the training focuses on learning the features to differentiate between the very similar classes.
An interesting question here is just how well the object detector works when the classes are so similar.
In an image classifier setting, it would make sense that the model could learn the subtle features that are unique to each of the classes.
As we previously covered, the object detector is a multi-task problem. 
It follows that the network has to both decide the anchors most likely containing an object and also determining the class of the object.
If we look at the classes for the COCO dataset for object detection, for example, we can see that there are 80 classes, but they are relatively difficult to mix up due to being quite distinct from one another. 
In Figure 6.6, we have an example of two images from the pets dataset showing two classes that would be relatively easy to mix up.

@@

Since we have not talked about image segmentation, we will briefly cover what it is about.
Image segmentation is also a very popular and important task with applications in many domains.
The goal for an image segmentation model is to determine the meaning of an image on a per-pixel level.
For example, in our case, we are going to train a model on the road segmentation data where the goal is to predict where the current lane goes, so we would like to know where the pixels that represent the current lane the car is on are.
So the output is a mask telling what class each pixel in the image belongs to.

@@

There is no complete implementation of the EfficientDet image segmentation head, and the authors don't really describe it too much in the paper.
To create the model, we took the EfficientDet detector as a base and augmented it with a segmentation head.
The segmentation head was based on the implementations of BiFPN segmentation from \citep{segment_pytorch}.
After creating multiple multi-task models, the creation and modification of the training loop were no longer so difficult to adjust to the new task.
Figuring out exactly how to add the new task into the existing model architecture took a while to find the exact values that the segmentation model needs for creating outputs.
After configuring data loaders and the model, the training was pretty much along the lines of the previous models we have covered.
Our modified EfficientDet architecture is shown in Figure 6.7.

@@

With the segmentation head and object detector head configured, it would now be easy to extend this model further.
We could easily add a new image classification head using the same backbone, like \citep{multinet}, where a single model is used for road segmentation, car detection, and road type classification.
Adding more classes to the segmentation or to the object classifier would be relatively easy to do.
Of course, there is a lot of sharing optimization to be done as well.
In our implementation, we naively shared everything, which was relatively detrimental to performance.
It would make sense to at least have the BiFPN be task-specific.
And since everything is now configured in the same model, it would be relatively easy to chop half of the backbone for each task and possibly get better results.
As we have covered over and over again, the space for possible improvements is huge.

\section{Recap}

>>>>

The ibean dataset \citep{beansdata} contains 1296 images classified into three classes; Healthy, Angular Leaf Spot, and Bean Rust.
The data set provides a split for the images that will be the split that is used for the experiments.
The citrus leaves dataset \citep{citrusdata}, on the other hand, contains 759 examples of citrus fruits and leaves with six different classes; Black Spot, Canker, Greening, Scab, Melanose, and Healthy.
This set is quite tricky as it contains 609 images of leaves and only 150 images of actual citrus fruits, and some of the classes have very few images.
Some examples from the data sets are in Figure 6.1.
For the training split, the data will be split in 65/10/25 ratio for the training, validation, and test sets, respectively.
The goal of this kind of split is to have some relatively reliable test results by having proportionally larger than regular test set since there is so little data.
Both of these datasets have a relatively small amount of images available for training, so it should serve as a good basis for trying to train multi-task classifiers and give a glimpse into the adjustments that need to be made when training multi-task classifiers.
The goal for this experiment is to get familiar with the basics of how to create the multi-task models, and how to do the actual changes to the training loop in the Pytorch code to accommodate for the multi-task training.

When the two closely related datasets are used to train the model, we could expect the model to generalize better on both of them, as hopefully similar features would benefit both tasks.
But as the tasks are quite simple, the benefits can be relatively small as the likely accuracy for the single-task models should be quite high.
In the case of the citrus data set, learning to classify the citrus fruits could prove quite difficult, and for that, the bean leaves likely won't be of much help.
This model is depicted in figure 6.2.

@@

Training both tasks in a multi-task setting, we found that the accuracy on the citrus data set was increased, and ibean accuracy stayed the same.
Specifically, the accuracy for the EfficientNet classifier went up to 88\% and for ResNet up to 89\%.
The final model evaluated on the test set was the model with the highest average validation accuracy over the two tasks when trained for 25 epochs.
All the results are displayed in table 6.1.
The results show that even though these tasks were relatively simple and the accuracies for the single-task models were high, there was a benefit to be gained from multi-task training.
Here we mainly focused on getting good results on a fully shared backbone, and more experiments could be done by having fewer parameters shared between the tasks.
As an introduction to multi-task learning, the experiments show that picking the correct sampling ratio is extremely important for achieving desired results.
In that way, the sampling ratio seems quite unlike learning rate in that picking too small learning rate may lead to slow convergence but will most likely produce results in the end.
With the sampling ratio, it seems that the correct value needs to be found by experimentation, as a too low ratio for the citrus data gives accuracies that can be as low as 60\%.

@@

The second way of approaching this problem is to first train a model on the COCO classes and to create a detector for the cat and dog classes.
Using the high-level model, we can automatically create acceptable object detection labels for the 37 classes in the pets dataset.
Then we can initialize a 37 class object detector using the weights of the two-class model.
By having such a well-initialized model for the detection, the second part of the training focuses on learning the features to differentiate between the very similar classes.
An interesting question here is just how well the object detector works when the classes are so similar.
In an image classifier setting, it would make sense that the model could learn the subtle features that are unique to each of the classes.
As we previously covered, the object detector is a multi-task problem. 
It follows that the network has to both decide the anchors most likely containing an object and also determining the class of the object.
If we look at the classes for the COCO dataset for object detection, for example, we can see that there are 80 classes, but they are relatively difficult to mix up due to being quite distinct from one another. 
In Figure 6.5, we have an example of two images from the pets dataset showing two classes that would be relatively easy to mix up.

@@

Since we have not talked about image segmentation, we will briefly cover what it is about.
Image segmentation is also a very popular and important task with applications in many domains.
The goal for an image segmentation model is to determine the meaning of an image on a per-pixel level.
For example, in our case, we are going to train a model on the road segmentation data where the goal is to predict where the current lane goes, so we would like to know where the pixels that represent the current lane the car is on are.
So the output is a mask telling what class each pixel in the image belongs to.
An example image and its segmentation mask are shown in Figure 6.6.

@@

There is no complete implementation of the EfficientDet image segmentation head, and the authors don't really describe it too much in the paper.
To create the model, we took the EfficientDet detector as a base and augmented it with a segmentation head.
The segmentation head was based on the implementations of BiFPN segmentation from \citep{segment_pytorch}.
After creating multiple multi-task models, the creation and modification of the training loop were no longer so difficult to adjust to the new task.
Figuring out exactly how to add the new task into the existing model architecture took a while to find the exact values that the segmentation model needs for creating outputs.
After configuring data loaders and the model, the training was pretty much along the lines of the previous models we have covered.
The results for our models were along the lines that we have seen previously, the multi-task model being a few points below the single-task models.
It could be possible here that the segmentation model could get better if it had its own BiFPN network and could only share the backbone with the detector, but we did not have enough time to run all possible experiments.
Our modified EfficientDet architecture is shown in Figure 6.7.

@@

With the segmentation head and object detector head configured, it would now be easy to extend this model further.
We could easily add a new image classification head using the same backbone, like \citep{multinet}, where a single model is used for road segmentation, car detection, and road type classification.
Adding more classes to the segmentation or to the object classifier would be relatively easy to do.
Of course, there is a lot of sharing optimization to be done as well.
In our implementation, we naively shared everything, which was likely relatively detrimental to performance.
It would make sense to at least have the BiFPN be task-specific.
And since everything is now configured in the same model, it would be relatively easy to chop half of the backbone for each task and possibly get better results.
As we have covered over and over again, the space for possible improvements is huge.

\section{Overview of the experiments}
Based on our experiments, we can say that multi-task learning is hard when compared to training single-task models.
Training multiple single-task models is easier and faster and requires less user input and parameter optimization.
The effect of many parameters, like the learning rate, can be evaluated relatively easily based on a few epochs in a single task setting.
Perhaps to find the very best model, more experimentation and exploration is needed, but just some default values can generally provide good results.

For someone with little experience in training multi-task models, it can be difficult to get started.
Whereas it is easy to find multiple examples of just how a training loop should be created, finding such examples for multi-task models is much more difficult.
For experimenting with different ways of splitting the features into local and global features, it is important to have systems for easily doing experiments on different architectures.
In our experiments, we created four points where it was possible to cut the last block in the ResNet, for example, so it was easy to create models that each had none to some local features.
Devising ways to split on a more fine-grained level and grouping tasks when there are multiple would be extremely important when searching for the optimal architecture.
Manually changing how the weights are shared would not be enough.
In our experiments, this was also quite clear, as we did not experiment beyond having models that have some or none local features.
We did not do any experiments where we would have grouped the features in a more well thought out way.
Likely, most benefits are gained when multiple tasks share half of the networks instead of only having a head be local.
This would make sense as the lower level features are more general and likely there would not be as many conflicts as in the deeper layers.
The number of things to tune is also quite discouraging when the model does not train properly.
Finding out which part of the modifications to the default process is wrong can be extremely difficult.
Some of the mistakes we encountered were difficult and frustrating to solve. 
For example, we forgot to do a deep copy of the local features and just passed the same features as references to all tasks and wondered why the training of a single task was impacting others so much.
Or when we froze all local features except the head when we meant only to freeze all the global features.
Or mixing up the order of the tasks in the model task heads, so the final inferences ended up being complete garbage.

The impact of the various parameters that have to be picked seemed to be much larger.
With an incorrect sampling ratio, the models could be extremely far from the good sampling ratios.
It seems that finding the perfect sampling ratio might not be possible, and settling for a good enough one might be a good idea.
It seems that it may be a good idea to pick a lower learning rate than normal to find the best intersection of features.

Finally, the model structure's changes to create the various classifiers' heads can be difficult to add.
While adding multiple image classifiers is relatively trivial to do if one is comfortable doing basic model definitions, it can be difficult to combine the various task heads for the more complex tasks.
For example, object detector models are generally provided with very specific training loops, so getting an object detector that has a classifier head is much more difficult than just training the two separately.
Without some easy way of abstracting the classifier heads and automatically injecting them into the training loop, it may not be possible for some to create multi-task models without extensive help.
We considered relatively simple tasks that already had open-source implementations and mostly combined them into a single model.
When dealing with some more advanced models like some of the ones presented in the multi-tasking chapter, it would be even more difficult to create the models.

Despite all the challenges, we did manage to create models that came close or were equal to their single-task counterparts.
The multi-task models were smaller and faster in those cases, confirming that using the shared representations can be an excellent idea.
